---
title: "Reproducible Research Assignment1"
author: "Karthik R"
date: "Monday, June 08, 2015"
output: html_document
---
Read raw data
```{r, echo=TRUE}
RawData <- read.csv("activity.csv")
library("data.table")
DT <- data.table(RawData)
```
Total Steps Taken per day
```{r, echo=TRUE}
DTS<-DT[, sum(steps, na.rm=TRUE), by = list(date)]
setnames(DTS,"V1","Total Steps")
DTS
```
Histogram of steps taken per day
```{r, echo=TRUE}
hist(DTS$'Total Steps', main="Steps taken per Day", ylab = "Number of Days", xlab="Steps",ylim=c(0,30),col=terrain.colors(7))
```

Mean and Median Steps taken per day
```{r, echo=TRUE}
DT[, list(Mean=mean(steps, na.rm=TRUE),Median=as.double(median(as.double(steps), na.rm=TRUE))), by = list(date)]
```
Calculate Average number of steps taken by interval
```{r, echo=TRUE}
DTSS<-DT[, mean(steps,na.rm=TRUE),by = list(interval)]
setnames(DTSS,"V1","Average Steps")
plot(DTSS$interval,DTSS$'Average Steps', type="l"
     ,ylab="Average Steps",xlab="Interval")
```

Maximum step is taken at the interval
```{r, echo=TRUE}
DTSS['Average Steps'==max(DTSS$'Average Steps'),]
```

Total data rows with "NA"
```{r, echo=TRUE}
sum(complete.cases(DT))
```

Fill the missing values using the Average value of steps by interval
```{r, echo=TRUE}
DTM<-merge(DT,DTSS, by="interval")
DTAdjusted<-DTM[,AdjustedSteps:=ifelse(is.na(DTM$steps),DTM$'Average Steps',DTM$steps)]
```

Calculate Total Steps Taken per day using adjusted data set
```{r, echo=TRUE}
DTSAdjusted<-DTAdjusted[, sum(AdjustedSteps), by = list(date)]
setnames(DTSAdjusted,"V1","Total Steps")
DTSAdjusted
```

Histogram of steps taken per day using Adjusted Data
```{r, echo=TRUE}
hist(DTSAdjusted$'Total Steps', main="Steps taken per Day", ylab = "Number of Days", xlab="Steps",ylim=c(0,30),col=terrain.colors(7))
```

Mean and Median Steps taken per day on adjusted data
```{r, echo=TRUE}
DTAdjusted[, list(Mean=mean(AdjustedSteps),Median=as.double(median(as.double(AdjustedSteps)))), by = list(date)]
```

Comparing the Adjusted data and original data, mean and median for the days where NA is present for steps varies. Ex. 2012-10-01. Histogram could also be verified to see the steps taken during first 5 days reduced.

Convert the Factor value date to Date Time
```{r, echo=TRUE}
str(DTAdjusted)
DTAdjusted<-DTAdjusted[,date:=as.Date(date)]
str(DTAdjusted)
```

Add variable to qualify a day as weekday or weekend
```{r, echo=TRUE}
DTAdjusted<-DTAdjusted[,Day:=ifelse(weekdays(DTAdjusted$date) %in% c("Saturday","Sunday"),"weekend","weekday")]
head(DTAdjusted)
```

Steps by weekday and weekend

```{r, echo=TRUE}
DTSSAdjusted<-DTAdjusted[, mean(AdjustedSteps),by = list(interval,Day)]
setnames(DTSSAdjusted,"V1","AverageSteps")
library("ggplot2")
g<-ggplot(DTSSAdjusted,aes(interval,AverageSteps))
p<-g+facet_wrap(~Day,ncol=1)
p<-p+geom_line(size=0.8,color="skyblue") 
p<-p+theme(panel.background=element_rect(fill="white", colour="black"))
p<-p+theme(panel.grid=element_line(colour="white"))
p<-p+theme(strip.background=element_rect(fill="peachpuff", colour="black"))
p<-p+ylab("Number of steps")
p<-p+xlab("Interval")
print(p)
```

